---
title: "Data EDA"
author: "keerthiraj"
date: "11/11/2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---


# Statistical Machine Learning Project (Fall 2019)

## Exploratory data analysis

### Initial setup

```{r setup, include=FALSE}

library(dplyr)
library(readr)
library(ggplot2)
library(dummies)
library(fastDummies)

data <- read_csv("Datasets/black friday sales/train.csv", na = "NA")

```

### Data summary

```{r}

summary(data)

```




### A look at the target response for first 100 data samples

```{r}

data.pur <- data$Purchase[1:100]
plot(data.pur, type = 'b')
```



### Purchase Amount histogram

```{r}

ggplot(data) + geom_bar(mapping = aes(Purchase))

```


### Counting data entries for each Product ID

```{r}
prod_count <- data %>% 
  group_by(Product_ID) %>%
  count()

```


### Selecting Product ID with >1000 data entries

```{r}

selected.prod <- prod_count %>% 

  arrange(desc(n)) %>%
  
  filter(n > 1400)

ggplot(data = selected.prod) + geom_point(aes(seq(1:nrow(selected.prod)), n))
```


### Average Purchase amount Vs various features 

```{r}

data %>% 
  
  group_by(Gender) %>%
  summarise(avg_purchase = mean(Purchase, na.rm = TRUE))


```


```{r}

data %>% 
  group_by(Age) %>%
  summarise(avg_purchase = mean(Purchase, na.rm = TRUE))

```

```{r}

data %>% 
  group_by(Marital_Status) %>%
  summarise(avg_purchase = mean(Purchase, na.rm = TRUE))

```


```{r}

data %>% 
  group_by(Stay_In_Current_City_Years) %>%
  summarise(avg_purchase = mean(Purchase, na.rm = TRUE))

```

```{r}
data %>% 
  group_by(Occupation) %>%
  summarise(avg_purchase = mean(Purchase, na.rm = TRUE))

```



### Converting Age groups and Stay in current city (years) into numerical features

```{r}

data <- subset(data, Product_ID %in% selected.prod$Product_ID)

data <- subset(data, !Product_Category_1 %in% c(4, 12, 13, 19, 20))


# converting age variable to numeric
data$Age[data$Age == "0-17"] <- "15"
data$Age[data$Age == "18-25"] <- "21"
data$Age[data$Age == "26-35"] <- "30"
data$Age[data$Age == "36-45"] <- "40"
data$Age[data$Age == "46-50"] <- "48"
data$Age[data$Age == "51-55"] <- "53"
data$Age[data$Age == "55+"] <- "60"

data$Age <- as.integer(data$Age)


data$Stay_In_Current_City_Years[data$Stay_In_Current_City_Years == "4+"] <- "4"
data$Stay_In_Current_City_Years <- as.integer(data$Stay_In_Current_City_Years)

data$Gender[data$Gender == "F"] <- 0
data$Gender[data$Gender == "M"] <- 1

data$City_Category_A <- ifelse(data$City_Category == "A", 1, 0)
data$City_Category_B <- ifelse(data$City_Category == "B", 1, 0)
data$City_Category_C <- ifelse(data$City_Category == "C", 1, 0)

data$City_Category_A <- as.factor(data$City_Category_A)
data$City_Category_B <- as.factor(data$City_Category_B)
data$City_Category_C <- as.factor(data$City_Category_C)

data$Product_Category_2[is.na(data$Product_Category_2)] <- 0
data$Product_Category_3[is.na(data$Product_Category_3)] <- 0

data.final <- data[,!(names(data) %in% c("City_Category"))]

summary(data.final)
```



```{r}

write.csv(data.final, 'cleaned_data.csv')

```




